\calculusName{Hilbert's Axiomatic Calculus}
\calculusAcronym{}
\calculusLogic{Classical Logic}
\calculusLogicOrder{First-Order}
\calculusType{Axiomatic}
\calculusYear{1917}

\calculusAuthor{David Hilbert}

\entryTitle{Hilbert's Axiomatic Calculus}
\entryAuthor{Richard Zach}     

\maketitle

\begin{entry}{Hilbert}  

\begin{calculus}
Axioms:
\begin{align*}
\text{I.\quad} 1)\quad & X \lor X \rightarrow X &
\text{II.}\quad 1)\quad & (x)Z \rightarrow Z\\
2) \quad & X \rightarrow X \lor Y &
2) \quad & (x)F(x) \rightarrow (Ex)F(x)\\
3) \quad & X \lor Y \rightarrow Y \lor X &
3) \quad & (x)(Z \lor F(x)) \rightarrow ((x)Z \lor (x)F(x))\\
4) \quad & X \lor (Y \lor Z) \rightarrow (X \lor Y) \lor Z &
4) \quad & (x)(F(x) \rightarrow G(x)) \rightarrow \\
&&& \qquad ((x)F(x) \rightarrow (x)G(x))\\
5) \quad & (X \rightarrow Y) \rightarrow (Z \lor X \rightarrow Z \lor Y) &
5) \quad & (x)(y)F(x, y) \rightarrow (y)(x)F(x, y)\\
&& 6) \quad & (x)(y)F(x, y) \rightarrow (x)F(x, x)
\end{align*}
Rules:
\begin{enumerate}
\item Renaming of bound variables
\item Substitution for propositional and predicate variables
\item Universal instantiation, i.e.,
\[
\infer{\alpha(a)}{(x)\alpha(x)}
\]
\item Universal closure of additional argument places, i.e.,
\[
\infer{(x)\alpha(F(x))}{\alpha(X)}
\qquad
\infer{(y)\alpha(F(x_1, \dots, x_n, y))}{\alpha(F(x_1, \dots, x_n))}
\]
\item Modus ponens, i.e.,
\[
\infer{\beta}{\alpha & \alpha \rightarrow \beta}
\]
\end{enumerate}

\end{calculus}

\begin{clarifications}
  Hilbert used the symbol $\times$ for disjunction $\lor$, $+$ for
  conjunction $\land$, and overlining $\overline{\phantom{X}}$ for
  negation. $\alpha \rightarrow \beta$ is an abbreviation for
  $\overline{\alpha} \lor \beta$, and $\alpha \land \beta$ for
  $\overline{\overline{\alpha} \lor \overline{\beta}}$. The usual
  conventions about operator precedence apply. $(x)$ is a universal
  quantifier and $(Ex)$ an existential quantifier. The system is
  formulated in a language that distinguishes between propositional and
  predicate constants and variables. In particular, the axioms
  are \emph{not} understood as schemas. The substitution rule allows the
  replacement of propositional variables by a formula that contains no
  object variable free, and a predicate variable with $n$ arguments
  $x_1$, \dots, $x_n$ by an expression in which all and only 
  $x_1$, \dots, $x_n$ occur free.
\end{clarifications}

\begin{history}
  The system was first presented in lecture notes by Paul Bernays to Hilbert's
  course ``Principles of Mathematics'', taught in the Winter term 1917--18;
  see~\cite{Hilbert1917}.
  The system is based on the axiomatic proof system of Whitehead and Russell's
  \emph{Principia Mathematica},~\irefmissing{PrincipiaMathematica} but restricts
  the language to first order. It adds the explicit rules of substitution and
  renaming, and avoids the use of free variables in its axioms and theorems.
\end{history}


\end{entry}
