\calculusName{} 
\calculusAcronym{\LKmumutilde} 
\calculusLogic{Classical Logic} 
\calculusLogicOrder{First-Order}
\calculusType{Sequent Calculus} 
\calculusYear{2000} 

\calculusAuthor{Curien}
\calculusAuthor{Herbelin} 

\entryTitle{\LKmumutilde} 
\entryAuthor{Hugo Herbelin}  

\etag{Two-Sided Sequents}
\etag{List Cedents}
\etag{Term-annotated}

\maketitle

\begin{entry}{LKMuMuTilde}

\newcommand{\cut}[2]{\langle #1 | #2 \rangle}
\begin{calculus}

{\sc Structural subsystem}
\[
\infer[\mathit{Ax_R}]
      {\Gamma \vdash a:A ~|~ \Delta}
      {(a:A) \in \Gamma}
\qquad
\infer[\mathit{Cut}]
      {\cut{v}{e} : (\Gamma \vdash \Delta)}
      {\Gamma \vdash v:A ~|~ \Delta & \Gamma ~|~ e:A \vdash \Delta}
\qquad
\infer[\mathit{Ax_L}]
      {\Gamma ~|~ \alpha:A \vdash \Delta}
      {(\alpha:A) \in \Delta}
\]
\[
\infer[\mathit{Focus}_L]
      {\Gamma~|~ \tilde\mu a.c:A \vdash \Delta}
      {c : (\Gamma, a:A \vdash \Delta)}
\qquad
\infer[\mathit{Focus}_R]
      {\Gamma \vdash \mu \alpha.c:A ~|~ \Delta}
      {c : (\Gamma \vdash \alpha:A, \Delta)}
\]
{\sc Introduction rules}
\[
\infer[\wedge_L^i]
      {\Gamma ~|~ \pi_i\cdot e :  A_1 \wedge A_2 \vdash \Delta}
      {\Gamma ~|~ e : A_i \vdash \Delta}
\qquad
\infer[\wedge_R]
      {\Gamma \vdash (v_1,v_2) : A_1 \wedge A_2 ~|~ \Delta}
      {\Gamma \vdash v_1:A_1 ~|~ \Delta & \Gamma \vdash v_2 : A_2 ~|~ \Delta}
\]
\[
\infer[\vee_L]
      {\Gamma ~|~ [e_1,e_2] : A_1 \vee A_2 \vdash \Delta}
      {\Gamma ~|~ e_1:A_1 \vdash \Delta & \Gamma ~|~ e_2 : A_2 \vdash \Delta}
\qquad
\infer[\vee_R^i]
      {\Gamma \vdash \iota_i(v) :  A_1 \vee A_2 ~|~ \Delta}
      {\Gamma \vdash v : A_i ~|~ \Delta}
\]
\[
\infer[\rightarrow_L]
      {\Gamma ~|~ v\cdot e : A\rightarrow B \vdash \Delta}
      {\Gamma \vdash v:A ~|~ \Delta & \Gamma ~|~ e:B \vdash \Delta}
\qquad
\infer[\rightarrow_R]
      {\Gamma \vdash \lambda a.v : A \rightarrow B ~|~\Delta}
      {\Gamma, a:A \vdash v:B ~|~ \Delta}
\]
\[
\infer[\exists_L]
      {\Gamma ~|~ \tilde\lambda x.e : \exists x\,A[x] \vdash \Delta}
      {\Gamma ~|~ e: A[y] \vdash \Delta}
\qquad
\infer[\exists_R]
      {\Gamma \vdash t \cdot v  : \exists x\,A[x] ~|~ \Delta}
      {\Gamma \vdash v :  A[t] ~|~ \Delta}
\]
\[
\infer[\forall_L]
      {\Gamma ~|~ t \cdot e : \forall x\, A[x] \vdash \Delta}
      {\Gamma ~|~ e: A[t] \vdash \Delta}
\qquad
\infer[\forall_R]
      {\Gamma \vdash \lambda x.v  : \forall x\,A[x] ~|~ \Delta}
      {\Gamma \vdash v : A[y] ~|~ \Delta}
\]
\[
\infer[\bot_L]
      {\Gamma ~|~ [~] : \bot \vdash \Delta}
      {}
\qquad
\infer[\top_R]
      {\Gamma \vdash () : \top ~|~ \Delta}
      {}
\]

\end{calculus}

\begin{clarifications}
There are three kinds of sequents: first $\Gamma \vdash v:A ~|~
\Delta$ with a distinguished formula on the right for typing the
program $v$, second $\Gamma ~|~ e:A \vdash \Delta$ with a
distinguished formula on the left for typing the evaluation context
$e$, and finally $c : (\Gamma \vdash \Delta)$ with no distinguished
formula for typing command $c$, i.e. the interaction of a program
within an evaluation context. The typing contexts $\Gamma$ and
$\Delta$ are lists of named formulas so that a non-ambiguous
correspondence with $\lambda$-calculus is possible (if it were sets or
multisets, there were e.g. no way to distinguish the two distinct
proofs of $x:A, x:A \vdash x:A ~|~$). Weakening rules are implemented
implicitly at the level of axioms. Contraction rules are derived,
using a cut against an axiom. No exchange rule is needed.  Not all
cuts are eliminable: only those not involving an axiom rule are.
Negation $\neg A$ can be defined as $A \rightarrow \bot$. In the rules
$\exists_E$ and $\forall_R$, $y$ is assumed fresh in $\Gamma$,
$\Delta$ and $A[x]$. The syntax of the underlying $\lambda$-calculus
is:
$$
\begin{array}{lll}
c & ::= & \cut{v}{e}\\
e & ::= & \alpha ~|~ \tilde\mu a.c ~|~ \pi_i\cdot e ~|~ [e,e] ~|~ v \cdot e ~|~ (t,e) ~|~ \tilde\lambda x.e ~|~ []\\
v & ::= & a ~|~ \mu\alpha.c ~|~ (v,v) ~|~ \iota_i(v) ~|~ \lambda a.v ~|~ \lambda x.v ~|~ (t,v) ~|~ ()\\
\end{array}
$$
\end{clarifications}

\begin{history}
The purpose of this system is to provide with a
$\lambda$-calculus-style computational meaning to Gentzen's LK
\iref{GentzenLK} and to highlight how the symmetries of sequent
calculus show computationally. Seeing the rules as typing rules, the
left/right symmetry is a symmetry between programs and their
evaluation contexts. At the level of cut elimination, giving priority
to the left-hand side relates to call-by-name evaluation while giving
priority to the right-hand side relates to call-by-value
evaluation~\cite{CurienHerbelin00}. Thanks to the presence of two dual
axiom rules and implicit contraction rules, the system supports a
tree-like sequent-free presentation like originally presented by
Gentzen for natural deduction~\cite{HerbelinHdR} (see
\iref{LKMuMuTildeTree}). The system can be seen as a symmetric variant
of $\lambdabar$-calculus~\iref{LambdaBar}.

The structural subsystem can be adapted to various sequent
calculi. Restriction to intuitionistic logic can be obtained by
demanding that the right-hand side has exactly one formula.

The presentation of this calculus with conjunctive and disjunctive
additive connectives has been studied in
\cite{Wadler03,HerbelinHdR}. A variant with only commands,
called ${\cal X}$, has been studied in \cite{BakLenLes05}, based on
previous work in~\cite{UrbanPhD}.  Various extensions of the system
emphasizing different symmetries can be found in the literature.
\end{history}

\begin{technicalities}
The system is obviously logically equivalent to Gentzen's $\LK$ when
equipped with the corresponding connectives and observed through the
sequents of the form $\Gamma \vdash \Delta$. 
\end{technicalities}


\end{entry}
