
\preface

The \textbf{Encyclopedia of Proof Systems} aims at providing a reliable, technically accurate, historically informative, concise, uniform and convenient central repository of proof systems for various logics. The goal is to facilitate the exchange of information among logicians, in order to foster and accelerate the development of proof theory and automated deduction.

Preparatory work for the creation of the Encyclopedia, such as the implementation of the LaTeX template and the setup of the Github repository, started in October 2014, triggered by the call for workshop proposals for the 25th Conference on Automated Deduction (CADE). Christoph Benzm\"uller, CADE's conference chair, and Jasmin Blanchette, CADE's workshop co-chair, encouraged me to submit a workshop proposal and supported my alternative idea to organize instead a special poster session based on encyclopedia entries. I am thankful for their encouragement and support.

In December 2014, Bj\"orn Lellmann, Giselle Reis and Martin Riener kindly accepted my request to beta-test the template and the instructions I had created. They submitted the first few example entries to the encyclopedia and provided valuable feedback, for which I am grateful. Their comments were essential for improving the templates and instructions before the public announcement of the encyclopedia.

In July 2015, Julian R\"oder's assistance was essential for the successful organization of the poster session at CADE. Cezary Kaliszyk and Andrei Paskevitch kindly allowed me to organize a discussion session as part of the Proof Exchange for Theorem Proving (PxTP) workshop, where the participants provided useful feedback and many ideas for improvements. Discussions with Lev Beklemishev, Bj\"orn Lellmann, Tomer Libal, Roman Kuznets, Sergei Soloviev, Valeria de Paiva and Anna Zamansky also brainstormed many ideas for improving the organization and structure of the encyclopedia.

In the few months that preceded CADE, as many as 63 entries, spanning a wide range of deduction styles and logics, have been submitted by 34 contributors. Although large for a single event, these numbers are still small compared to the vast number of proof systems that have been invented and to the number of people who work on logical calculi nowadays. Therefore, this community-wide initiative to produce an encyclopedia is only at the beginning and will continue to be open to submissions for a long time. 

\vspace{\baselineskip}
\begin{flushright}\noindent
October 2016\hfill {\it Bruno Woltzenlogel Paleo}
\end{flushright}


